---
title: "Week 6"
output: html_document
date: "2024-06-30"
---

Read data of 1996:

```{r, echo=TRUE}
library(readr)
data <- read_csv("C:/Users/lyy03/OneDrive/Documents/Mcdaniel/ANA 515/Week 6/StormEvents_details-ftp_v1.0_d1996_c20220425.csv.gz")

```

Clean data:
1. select specific columns
2. sort the data by STATE
3. change state and county names to title case
```{r, echo=TRUE}
library(dplyr)
library(stringr)
data_sorted <- data %>% select(BEGIN_YEARMONTH, EPISODE_ID, STATE, STATE_FIPS, CZ_NAME, CZ_TYPE, CZ_FIPS, EVENT_TYPE) %>% arrange(STATE) %>% mutate (
  STATE = str_to_title(STATE),
  CZ_NAME = str_to_title(CZ_NAME)
)
```

Limit to the events listed by county FIPS (CZ_TYPE of “C”) and then remove the CZ_TYPE column :
```{r, echo = TRUE}
data_filtered <- data_sorted %>% 
    filter(CZ_TYPE == "C") %>%
    select(-CZ_TYPE
           )
```

Adjust state and county FIPS and create a new FIPS column:
```{r, echo = TRUE}
library(tidyr)
data_ncol <- data_filtered %>%
  mutate(
    STATE_FIPS = str_pad(STATE_FIPS, width = 2, side = "left", pad = "0"),
    CZ_FIPS = str_pad(CZ_FIPS, width = 3, side = "left", pad = "0") )%>%
      unite("FIPS", STATE_FIPS, CZ_FIPS, sep = "", remove = FALSE)
  
```

Rewrite column names:
```{r, echo=TRUE}
data_ncol <- data_ncol %>% rename_all(tolower)
```

Read state data from R
```{r, echo = TRUE}
data(state)
state_data <- data.frame(
  name = state.name,
  region = state.region,
  area = state.area
)
```

merge datasets and clean:
```{r, echo = TRUE}
event_counts <- data_ncol %>%
    group_by(state) %>%
    summarise(event_count = n())
state_data <- state_data %>% rename(state=name)
merged_data <- merge(state_data, event_counts, by = "state", all.x=FALSE)

```

Scatterplot of number of storms events in 1996 by land area by region:
```{r, echo = TRUE}
library(ggplot2)
ggplot(data = merged_data, aes(x=area, y=event_count, color = region)) +
  geom_point() +
  labs(
    x = "Land area (square miles)",
    y = "# of storms events in 1996",
    color = "region"
  ) +
  theme_minimal()+
  theme(
    legend.position = "right"
  )
```
